---
title: "RWorksheet_Octaviano#4b"
author: "Jirraine S. Octaviano"
date: "2023-11-08"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
Using Loop Function
for() loop

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
   Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r setup, include=FALSE}
vector_A <- c(1,2,3,4,5)

matrix_A <- matrix(0,nrow = 5, ncol =5)

for (j in 1:5)
for (k in 1:5)
{
    matrix_A[j,k] <- abs (vector_A[j] - vector_A[k])
}

  matrix_A
  
```
2. Print the string "*" using for() function. The output should be the same as shown in Figure
```{r setup, include=FALSE}
  for (j in 1:5) {
  cat(paste0("\"", rep("*", j), "\""), "\n")
  }

```
#3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r name, echo = TRUE}

userInput <- as.integer(readline("Enter starting number for Fibonacci sequence: "))

x <- userInput
y <- 0


cat("Fibonacci sequence starting from", userInput, ":\n")

repeat {
 
  next_num <- x + y
 
  if (next_num > 500){
    break
  }
  cat(next_num, " ")
  x <- y
  y <- next_num
 
}

```
Using Basic Graphics (plot(),barplot(),pie(),hist())

4. Import the dataset as shown in Figure 1 you have created previously.

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset?
Show your codes and its result.
```{r name, echo = TRUE}
prevData <- read.csv("HouseHold DATA.csv")
head(prevData)
```
b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r name, echo = TRUE}
males <- prevData[prevData$Gender == "M",]
males

females <- prevData[prevData$Gender == "F",]
females

numofMale <- nrow(males)
numofMale

numofFem <- nrow(females)
numofFem
```
c. Create a graph for the number of males and females for Household Data. Use plot(),chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and its result.
```{r name, echo = TRUE}

totalMaleFemale <- table(prevData$Gender)
barplot(totalMaleFemale,
        main = "Number of Males and Females",
        xlab = "Gender",
        ylab = "Count",
        col = c("red", "blue"),
        legend.text = rownames(totalMaleFemale),
        beside = TRUE)
```
5. The monthly income of Dela Cruz family was spent on the following:
5a. a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.

```{r name, echo = TRUE}

spending_data <- data.frame(
  Category = c("Food", "Electricity", "Savings", "Miscellaneous"),
  Value = c(60, 10, 5, 25)
)

spending_data$Percentage <- spending_data$Value / sum(spending_data$Value) * 100

colors <- c("red", "blue", "green", "purple")

pie(spending_data$Value,
    labels = paste(spending_data$Category, "(",spending_data$Percentage,"%)"),
    col = colors,
    main = "Monthly Income Spending of Dela Cruz Family")

legend("topright", spending_data$Category, fill = colors)


```
