---
title: "RWorksheet#4c"
author: Jirraine S. Octaviano
date: "2023-11-22"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Use the dataset mpg
• Download and open the mpg file. Upload it to your OWN environment

a. Show your solutions on how to import a csv file into the environment.

```{r}

  library(readr)
  mpg_file <- read.csv ("mpg.csv")
  mpg_file

```
b.Which variables from mpg dataset are categorical?

```{r}
is_Cont <- sapply(mpg, is.character)
Cont_Vars <- names(mpg)[is_Cont]
Cont_Vars

```
c. Which are continuous variables?

```{r}
  is_Cont <- sapply(mpg, is.numeric)
  Cont_Vars <- names(mpg)[is_Cont]
  Cont_Vars

```

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.

```{r}
manufacturer_AsTablE <- table(mpg$manufacturer)
manufacturer_most_models <- names(manufacturer_AsTablE )[which.max(manufacturer_AsTablE)]

manufacturer_most_models

model_AsTablE  <- table(mpg$model)
model_most_vars <- names(model_AsTablE )[which.max(model_AsTablE)]

model_most_vars
## The dodge manufacturer has the most models
## The caravan 2wd has the most variations

```
a. Group the manufacturers and find the unique models. Show your codes and result.

```{r}
#install.packages("dplyr")
library(dplyr)
ManuFacturer_Models <- data.frame(manufacturer = mpg$manufacturer, model = mpg$model)

unique_mods <- unique(ManuFacturer_Models)

unique_mods
unique_mods$mpg.model

```
b. Graph the result by using plot() and ggplot(). Write the codes and its result.

```{r}

library(ggplot2)
library(dplyr)

unique_plot <- plot(as.factor(factored$Manufacturer),
                    main = "Unique Models of Manufacturers",
                    xlab = "Manufacturer", 
                    ylab = "Number of Unique Models",
                    cex.names = 0.8, las = 2)



unique_count <- unique_mods %>% 
  count(unique_mods$manufacturer)
unique_count 

ggplot(unique_count, aes(x = manufacturer, y = n)) +
  geom_bar(stat = "identity", fill = "red") + 
  labs(x = "Manufacturer", y = "Number of Unique Models") +
  ggtitle("Unique Models of Manufacturers") +
  theme_minimal()

```
2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.
a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point() 

## In this plot, where each point corresponds to a specific car model and its manufacturer. 
## The x-axis will represent the car model, the y-axis will represent the manufacturer, and each point will indicate a specific combination of model and manufacturer in the dataset.
```
b.For you, is it useful? If not, how could you modify the data to make it more informative?

```{r}

##It is useful if you want to know how many models each manufacturer has. 
##Absolutely, your proposed modifications sound like excellent ways to make the plot more informative. Color coding based on the classvariable and incorporating labels with the cty variable.

```

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}

ggplot(head(mpg,20), aes(x = model, y = year)) + geom_point()

```
4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}

library(dplyr)

Model_car_Count <- mpg %>% 
  group_by(model) %>%
  summarize(Number_Of_Cars = n())

Model_car_Count

```

a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results

```{r}

obs_20 <- head(Model_car_Count, 20)

top_20 <- ggplot(obs_20, aes(x = model, y = Number_Of_Cars, fill = model)) + geom_bar(stat = "identity") + labs(Title = "Number Of Cars per Model", x = "Model" , y = "Number Of Cars")

top_20

```
b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.

```{r}

flipped_Top_20 <- ggplot(obs_20, aes(x = model, y = Number_Of_Cars, fill = model)) + geom_bar(stat = "identity") + labs(Title = "Number Of Cars per Model", x = "Model" , y = "Number of cars") + coord_flip()

flipped_Top_20
```
``` {r}
flipped_Top_20 <- ggplot(obs_20, aes(x = Model, y = Number_Of_Cars, fill = Model)) +
  geom_bar(stat = "identity") +
  labs(title = "Number Of Cars per Model", x = "Model", y = "Number of cars") +
  coord_flip()

flipped_Top_20
```

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

```{r}

Cyl_Displ_Plot <- ggplot(mpg, aes(x = cyl, y = displ, Color = displ)) +
  geom_point() +
  labs(Title = "Relationship Between No. Of Cylinders And Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement")

Cyl_Displ_Plot

```

a. How would you describe its relationship? Show the codes and its result.

```{r}
# It will generate a scatter plot showing the relationship between the number of cylinders and engine displacement. The color of each point represents the corresponding engine displacement. 

# As the number of cylinders increases, the engine displacement tends to increase as well. This suggests that there is a tendency for larger engines with more cylinders to have greater displacement.
##The scatter plot with a linear regression line and color-coded points illustrates the relationship between cylinder number and engine displacement. 
##It suggests that as cylinder number increases, engine displacement also increases, indicating larger engines with more cylinders have greater displacement.
```

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?

```{r}

displ_hwy_plot <- ggplot(mpg, aes(x = displ, y = hwy, Color = cty)) +
  geom_point() +
  labs(title = "Relationship Between Engine Displacement And Highway MPG",
       x = "Engine Displacement",
       y = "Highway MPG")
displ_hwy_plot

## This scatter plot visually illustrates the relationship between engine displacement (displ) and highway miles per gallon (hwy), utilizing color to represent city fuel economy (cty). The plot enables a visual analysis of how varying levels of city fuel economy correlate with engine displacement and highway miles per gallon, offering insights into the fuel efficiency of vehicles with different engine sizes.

```
6. Import the traffic.csv onto your R environment.

a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show your answer.

```{r}
library(readr)
traffic <- read.csv("Traffic.csv")

Num_obs <- nrow(traffic)
Num_obs

Num_Vars <- ncol(traffic)
Num_Vars

Vars <- colnames(traffic)
Vars

```

b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}

Junctions_subset_1 <- subset(traffic, Junction == 1)

Junctions_subset_2 <- subset(traffic, Junction == 2)

Junctions_subset_3 <- subset(traffic, Junction == 3)

Junctions_subset_4 <- subset(traffic, Junction == 4)

```
c. Plot each junction in a using geom_line(). Show your solution and output.
```{r}

junctions_1_plot <- ggplot(Junctions_subset_1, aes(x = as.Date(junctions_subset_1$DateTime), y = Vehicles)) + 
  geom_line() +
  scale_x_date(date_labels = "%b-%Y") + theme(legend.position = "none") +
  labs(title = "Junction 1", x = "Time", y = "Number of Vehicles")

junctions_1_plot

Junction_2_Plot <- ggPlot(junctions_subset_2, aes(x = as.Date(junctions_subset_2$DateTime), y = Vehicles)) + 
  geom_line() +
  Scale_x_Date(date_labels = "%b-%Y") + theme(legend.position = "none") +
  labs(Title = "Junction 2", x = "Time", y = "Number of Vehicles")

Junction_2_Plot

Junction_3_Plot <- ggPlot(junctions_subset_3, aes(x = as.Date(junctions_subset_3$DateTime), y = Vehicles)) + 
  geom_line() +
  Scale_x_Date(date_labels = "%b-%Y") + theme(legend.position = "none") +
  labs(Title = "Junction 3", x = "Time", y = "Number of Vehicles")

Junction_3_Plot

Junction_4_Plot <- ggPlot(junctions_subset_4, aes(x = as.Date(junctions_subset_4$DateTime), y = Vehicles)) + 
  geom_line() +
  Scale_x_Date(date_labels = "%b-%Y") + theme(legend.position = "none") +
  labs(Title = "Junction 4", x = "Time", y = "Number of Vehicles")

junction_4_Plot

```
7. From alexa_file.xlsx, import it to your environment

a. How many observations does alexa_file has? What about the number of columns? Show your solution and answer.

b. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.

c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.

d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.

e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer.
```