---
title: "RWorksheet_Octaviano#4c"
author: "Plotting the Graph using ggplot2"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Use the dataset mpg
• Download and open the mpg file. Upload it to your OWN environment

a. Show your solutions on how to import a csv file into the environment.

```{r}

library(readr)
mpg_file <- read.csv ("mpg.csv")
mpg_file

```
b.Which variables from mpg dataset are categorical?

```{r}

str(mpg_file)

##  manufacturer, model, trans, dry, fl, class variables are categorical

```
c. Which are continuous variables?

```{r}
is_cont <- sapply(mpg, is.numeric)
cont_vars <- names(mpg)[is_cont]
cont_vars

```

2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.

```{r}

manufacturer_AstablE <- table(mpg_file$manufacturer)
names(manufacturer_AstablE)[which.max(manufacturer_asTable)]

## dodge manufacturer has the most models

model_AstablE <- table(mpg_file$model)
model_AstablE
names(model_AstablE)[which.max(model_AstablE)]

## caravan 2wd has the most variations

```
a. Group the manufacturers and find the unique models. Show your codes and result.

```{r}

manufacturers_models <- data.frame(Manufacturer = mpg_file$manufacturer, Model = mpg_file$model)

unique_mods <- unique(manufacturers_models)

unique_mods
unique_mods$mpg_file.model

```

b. Graph the result by using plot() and ggplot(). Write the codes and its result.


3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.

```{r}

4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result

a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results.

b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.

a. How would you describe its relationship? Show the codes and its result.

6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?

6. Import the traffic.csv onto your R environment.

a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show your answer.

b. subset the traffic dataset into junctions. What is the R codes and its output?

c. Plot each junction in a using geom_line(). Show your solution and output.

7. From alexa_file.xlsx, import it to your environment

a. How many observations does alexa_file has? What about the number of columns? Show your solution and answer.

b. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.

c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer.

d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.

e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship. Show your solution and answer.
```